---
title: "Cat and Mouse. Understanding Defensive Reaction to Offensive Motion"
author: "Jack Sullivan"
date: "2025-01-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

*"[Offensive] motion is about gaining leverage on the defense. It gives the quarterback an understanding of what coverage he's facing.* - Barry Alvarez

```{r, echo = FALSE}
# plays %>% filter(possessionTeam == "BAL" & defensiveTeam == "MIA" & preSnapHomeScore == 7 & preSnapVisitorScore == 7)

# all_tracking_motion %>% filter(gameId == 2022091800 & playId == 1108)

source("animation.r")

animate_presnap(2, 2022091800, 1108, "purple", "aquamarine")
```

(describe play, potentially improve visualization with center of mass visualization. Connect the play to the insights we can gather using the model.)


## Defensive Positioning Metrics

(Describe the metrics we use to measure defensive movement as a result of offensive motion. Utilize graphs and images to convey the idea of perimeter/center of mass/area)






## Most impactful Motion schemes

(Describe the motion types, formations for motion, ect. that are most impactful to defensive shifting)




## Coach's Application

(describe how our findings can be applied to game planning, play creation, in-game adjustments)


##
